TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-01-21 paper/draft.tex
make figures more readable

@@

The spectra contain roughly 8000 pixels,
so we might expect the $\chi^2$ values to be close to 8000 in magnitude.
They are much higher.
This discrepancy flows from the fact that---while the spectral model is good at the percent
level---the signal-to-noise of the typical spectra are more than 100.
That is, the $\chi^2$ values do show that the model is not good in the frequentist
sense; it is only good at the percent level.
This percent-level accuracy of the model in the data domain is also demonstrated...
\color{gcolor}{HOGG: Adam says he doesn't think this is right: if $\chi^2/n>>1$ then the model is NOT an accurate representation of the data.}\color{black} \\
\color{gcolor}{BIRKY: Can you demonstrate this by showing the RMS residuals, putting it in as a figure, and putting this paragraph back to me? I am imagining a figure that shows a scatter plot of T vs RMS residual, colored by metallicity. Or something like that.}\color{black}

@@

\subsection{Radii}
Using empirically calibrated relations from \citealt{Mann:2015} we determine radii... some diagnostics...

@@

As a validation test of our derived temperatures, we perform a comparison between several color-temperature relations from literature, which use combination of \zmass\ and visual band photometries to predict temperatures. To obtain visual band magnitudes for a set of sources, we cross-matched the 10,311 sources in our ``safe'' test sample to the AAVSO Photometric All-Sky Survey DR9 (\textsl{APASS}; \citealt{Henden:2016}), to obtain a subsample of 6,257 sources with both $BV$ photometries measured by \textsl{APASS}, and $JHK$ photometries from \apogee. Figure \ref{fig:teff_comparisons} shows \cannon\ vs. photometric temperatures on the right, and \aspcap\ vs. photometric temperatures on the left for each of the 6,257 sources, colored by their respective spectroscopic metallicities.

% SUMMARIZE \citealt{Casagrande:2008}, \citealt{Boyajian:2012}, and \citealt{Mann:2015} 
Shown in the top panels of \ref{fig:teff_comparisons}

We find \aspcap\ overestimates \teff\ on average by $130-170$K, while \thecannon\ underestimates \teff\ by $\sim25$K . We also see the scatter is much higher for the \aspcap\ temperatures ($\sim230$K) compared to the scatter of \cannon\ temperatures ($\sim85$K).

% The scatter of each relation being largely due to metallicity variation, both for the \cannon\ and \aspcap\ metallicities.
\cannon\ temperatures show the highest agreement with the photometric relations from \citealt{Mann:2015} and \citealt{Boyajian:2012}, with a scatter of 80K.

While the large bulk of points falling within the temperature ranges of the plots shown does not qualitatively show a lot more deviation from color-temperature relations, the average \aspcap\ temperatures are much hotter, and scatters are much higher skewed by 500 sources which \aspcap\ predicts to have temperatures over 4200K.

\subsection{Isochrone Comparisons}

@@

%%==========================================================================================
\section{Discussion} \label{sec:discussion}

\begin{enumerate}
\item[-] Brief summary of what we did and found

\item[-] Things we learned along the way: For example, not all training sets worked well in our experiments, probably because they were too small or had too little dynamic range.

\item[-] \color{gcolor}{HOGG: Some discussion of precision and accuracy: \thecannon\ can produce precise results, but is only as accurate as its training set.}\color{black}

\item[-] Some discussion of whether the derivatives are sensible or surprising. \color{gcolor}{HOGG: I'm not sure what to say here}\color{black}

\item[-] \color{gcolor}{HOGG:\thecannon\ as a test of evaluating training parameter accuracy?}\color{black}

\item[-] \color{gcolor}{HOGG: Discussion of validation tests and ground truth}\color{black}

\end{enumerate}

Future work:
\begin{enumerate}
\item[-] Create a binary sample; using color/magnitude \citealt{Widmark:2018} and by spectroscopy \citealt{ElBadry:2018}

\item[-] \color{gcolor}{HOGG: Discussion of the implications of these results for the NASA \tess\ Mission and how the future might change..?}\color{black}
\end{enumerate}

@@

%---------------------------
\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/demo_derivatives_teff1.png}
\includegraphics[width=16cm]{figures/demo_derivatives_teff2.png}
\includegraphics[width=16cm]{figures/demo_derivatives_teff3.png}
\end{center}
\caption{\textit{Top two panels of each plot:} Mann-trained model for varying temperatures, and similar metallcities. \textit{Third panel of each plot:} Derivative of \thecannon\ model with respect to temperature, taken at the median training temperature, T$_{\rm eff}=3463$K; error estimate computed using a jackknife statistic at each pixel is marked in red, making it possible to distinguish which features vary significantly with change in spectral type, and which are likely due to noise.}\label{fig:demo_teff}
\end{figure}

%---------------------------
\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/demo_derivatives_feh1.png}
\includegraphics[width=16cm]{figures/demo_derivatives_feh2.png}
\includegraphics[width=16cm]{figures/demo_derivatives_feh3.png}
\end{center}
\caption{\textit{Top two panels of each plot:} Mann-trained model for varying metallicities and similar temperatures. \textit{Third panel:} Derivative of \thecannon\ model with respect to metallicity, taken at the median training metallicity, [Fe/H]$=-0.03$ dex; jackknife computed error at each pixel is shown in red.} \label{fig:demo_feh}
\end{figure}

>>>>

% \usepackage{draftwatermark}
% \SetWatermarkScale{3}
% \SetWatermarkLightness{.7}

@@

The spectra contain roughly 8000 pixels,
so we might expect the $\chi^2$ values to be close to 8000 in magnitude.
They are much higher.
This discrepancy flows from the fact that---while the spectral model is good at the percent
level---the signal-to-noise of the typical spectra are more than 100.
That is, the $\chi^2$ values do show that the model is not good in the frequentist
sense; it is only good at the percent level.
This percent-level accuracy of the model in the data domain is also demonstrated...
% \color{gcolor}{HOGG: Adam says he doesn't think this is right: if $\chi^2/n>>1$ then the model is NOT an accurate representation of the data.}\color{black} \\
\color{gcolor}{BIRKY: Can you demonstrate this by showing the RMS residuals, putting it in as a figure, and putting this paragraph back to me? I am imagining a figure that shows a scatter plot of T vs RMS residual, colored by metallicity. Or something like that.}\color{black}

@@

\subsection{Radii}
% Using empirically calibrated relations from \citealt{Mann:2015} we determine radii... some diagnostics...
TBD

@@

TBD
% As a validation test of our derived temperatures, we perform a comparison between several color-temperature relations from literature, which use combination of \zmass\ and visual band photometries to predict temperatures. To obtain visual band magnitudes for a set of sources, we cross-matched the 10,311 sources in our ``safe'' test sample to the AAVSO Photometric All-Sky Survey DR9 (\textsl{APASS}; \citealt{Henden:2016}), to obtain a subsample of 6,257 sources with both $BV$ photometries measured by \textsl{APASS}, and $JHK$ photometries from \apogee. Figure \ref{fig:teff_comparisons} shows \cannon\ vs. photometric temperatures on the right, and \aspcap\ vs. photometric temperatures on the left for each of the 6,257 sources, colored by their respective spectroscopic metallicities.

% % SUMMARIZE \citealt{Casagrande:2008}, \citealt{Boyajian:2012}, and \citealt{Mann:2015} 
% Shown in the top panels of \ref{fig:teff_comparisons}

% We find \aspcap\ overestimates \teff\ on average by $130-170$K, while \thecannon\ underestimates \teff\ by $\sim25$K . We also see the scatter is much higher for the \aspcap\ temperatures ($\sim230$K) compared to the scatter of \cannon\ temperatures ($\sim85$K).

% % The scatter of each relation being largely due to metallicity variation, both for the \cannon\ and \aspcap\ metallicities.
% \cannon\ temperatures show the highest agreement with the photometric relations from \citealt{Mann:2015} and \citealt{Boyajian:2012}, with a scatter of 80K.

% While the large bulk of points falling within the temperature ranges of the plots shown does not qualitatively show a lot more deviation from color-temperature relations, the average \aspcap\ temperatures are much hotter, and scatters are much higher skewed by 500 sources which \aspcap\ predicts to have temperatures over 4200K.

\subsection{Isochrone Comparisons}
TBD

@@

%%==========================================================================================
\section{Discussion} \label{sec:discussion}
TBD

% \begin{enumerate}
% \item[-] Brief summary of what we did and found

% \item[-] Things we learned along the way: For example, not all training sets worked well in our experiments, probably because they were too small or had too little dynamic range.

% \item[-] \color{gcolor}{HOGG: Some discussion of precision and accuracy: \thecannon\ can produce precise results, but is only as accurate as its training set.}\color{black}

% \item[-] Some discussion of whether the derivatives are sensible or surprising. \color{gcolor}{HOGG: I'm not sure what to say here}\color{black}

% \item[-] \color{gcolor}{HOGG:\thecannon\ as a test of evaluating training parameter accuracy?}\color{black}

% \item[-] \color{gcolor}{HOGG: Discussion of validation tests and ground truth}\color{black}

% \end{enumerate}

% Future work:
% \begin{enumerate}
% \item[-] Create a binary sample; using color/magnitude \citealt{Widmark:2018} and by spectroscopy \citealt{ElBadry:2018}

% \item[-] \color{gcolor}{HOGG: Discussion of the implications of these results for the NASA \tess\ Mission and how the future might change..?}\color{black}
% \end{enumerate}

@@

%---------------------------
\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/demo_derivatives_teff1.png}
\includegraphics[width=16cm]{figures/demo_derivatives_teff2.png}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/demo_derivatives_teff3.png}
\end{center}
\caption{\textit{Top two panels of each plot:} Mann-trained model for varying temperatures, and similar metallcities. \textit{Third panel of each plot:} Derivative of \thecannon\ model with respect to temperature, taken at the median training temperature, T$_{\rm eff}=3463$K; error estimate computed using a jackknife statistic at each pixel is marked in red, making it possible to distinguish which features vary significantly with change in spectral type, and which are likely due to noise.}\label{fig:demo_teff}
\end{figure}

%---------------------------
\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/demo_derivatives_feh1.png}
\includegraphics[width=16cm]{figures/demo_derivatives_feh2.png}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/demo_derivatives_feh3.png}
\end{center}
\caption{\textit{Top two panels of each plot:} Mann-trained model for varying metallicities and similar temperatures. \textit{Third panel:} Derivative of \thecannon\ model with respect to metallicity, taken at the median training metallicity, [Fe/H]$=-0.03$ dex; jackknife computed error at each pixel is shown in red.} \label{fig:demo_feh}
\end{figure}
